Adaption of the Chumbley Score to matching of bullet striation marks

### 2018-04-13 doc/paper.Rmd
including locations

\tableofcontents
\newpage
```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  error = FALSE, 
  collapse = TRUE,
  comment = "",
  fig.height = 6,
  fig.width = 6,
  fig.align = "center",
  out.width= '\\textwidth',
  cache = FALSE,
  fig.path='figures/',
  echo=FALSE,
  cache=TRUE
)
options(knitr.table.format = "latex")
library(tidyverse)
library(kableExtra)
library(xtable)
```

@@

## Potential Challenges in Chumbley Score Adaptation

<!--Markings made by screw driver tips \citep{chumbley} are longer and pronounced and about 7 mm in length \citep{manytoolmarks1}, while the groove to groove length of the bullet lands from the Hamby study \citep{hamby} are about 2mm in length. The Ruger pistol barrels used in the  Hamby study have a 9mm caliber. This means that bullets used in these pistols have land markings of only about a quarter of the size of screwdriver toolmarks used by \citet{chumbley, hadler}, therefore raising the problem of potential loss in power of the test. -->

\hh{The Chumbley score allows a separation of a toolmark into raw and normalized digitized versions. Originally, this mechanism is intended to separate between class characteristics and individual markings, however, in the setting of matching bullet striae, we could also use it to separate bullet curvature in profiles from signatures before matching signatures. }

@@

In a pre-processing step  the two markings are smoothed using a lowess \citep{lowess} with coarseness parameter $c$. The purpose of this smoothing is to remove drift and (sub)class characteristics.

The Chumbley scores is calculated in two phases, namely, an optimization step and a validation step.
In the optimization step, the two markings are aligned horizontally such that within a pre-defined window of length $w_o$ the correlation between $x(t_1)$ and $y(t_2)$ is maximized:
\[
\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1, 1 \le t_2 \le T_2} \text{cor} \left(x (t_1, t_1 + w_o), y(t_2, t_2 + w_o) \right)
\]
This results in an optimal vertical (in-phase) shift of $t_1^o - t_2^o$ for aligning the two markings.

@@

library(gridExtra)
grid.arrange(p1,p2, ncol=2)
```

@@

```{r profile-sketch, fig.height = 2.5, fig.width='\\textwidth', fig.cap="Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \\ref{fig:sigs-profiles}. "}
profile <- sigs_graphics[[1]] %>% filter(land_id==19)
profile$lw25 <- lowess(x = profile$y, y = profile$value, f=.25)$y
profile$lw50 <- lowess(x = profile$y, y = profile$value, f=.5)$y
profile$lw75 <- lowess(x = profile$y, y = profile$value, f=.75)$y
profile$lw100 <- lowess(x = profile$y, y = profile$value, f=1)$y

profiles <- tidyr::gather(profile, coarseness, values, lw25:lw75)
profiles <- profiles %>% mutate(
  c = readr::parse_number(coarseness)/100
)
profiles %>% 
  ggplot(aes(x = y, y = value-values)) + geom_line() +
  theme_bw() +
  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) +
  facet_grid(~c, labeller="label_both")

>>>>

\tableofcontents
\newpage
```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  error = FALSE, 
  collapse = TRUE,
  comment = "",
  fig.height = 6,
  fig.width = 6,
  fig.align = "center",
  out.width= '\\textwidth',
  cache = FALSE,
  fig.path='figures/',
  echo=FALSE,
  cache=TRUE
)
options(knitr.table.format = "latex")
library(tidyverse)
library(kableExtra)
library(xtable)
library(gridExtra)
```

@@

<!--
## Potential Challenges in Chumbley Score Adaptation

Markings made by screw driver tips \citep{chumbley} are longer and pronounced and about 7 mm in length \citep{manytoolmarks1}, while the groove to groove length of the bullet lands from the Hamby study \citep{hamby} are about 2mm in length. The Ruger pistol barrels used in the  Hamby study have a 9mm caliber. This means that bullets used in these pistols have land markings of only about a quarter of the size of screwdriver toolmarks used by \citet{chumbley, hadler}, therefore raising the problem of potential loss in power of the test. -->

@@

In a pre-processing step  the two markings are smoothed using a lowess \citep{lowess} with coarseness parameter $c$. 
 Originally, this smoothing is intended to remove drift and (sub)class characteristics from individual markings, however, in the setting of matching bullet striae, we can also make use of this mechanism can be used to separate bullet curvature in profiles from signatures before matching signatures.

After removing sub-class structure, 
the Chumbley scores is calculated in two steps: an optimization step and a validation step.
In the optimization step, the two markings are aligned horizontally such that within a pre-defined window of length $w_o$ the correlation between $x(t_1)$ and $y(t_2)$ is maximized:
\[
\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1, 1 \le t_2 \le T_2} \text{cor} \left(x (t_1, t_1 + w_o), y(t_2, t_2 + w_o) \right)
\]
This results in an optimal vertical (in-phase) shift of $t_1^o - t_2^o$ for aligning the two markings.

@@

@@

```{r profile-sketch, fig.height = 5, fig.width='\\textwidth', fig.cap="Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \\ref{fig:sigs-profiles}. "}
profile <- sigs_graphics[[1]] %>% filter(land_id==19)
profile$lw20 <- lowess(x = profile$y, y = profile$value, f=.20)$y
profile$lw25 <- lowess(x = profile$y, y = profile$value, f=.25)$y
profile$lw30 <- lowess(x = profile$y, y = profile$value, f=.30)$y
#profile$lw50 <- lowess(x = profile$y, y = profile$value, f=.5)$y
#profile$lw75 <- lowess(x = profile$y, y = profile$value, f=.75)$y
#profile$lw100 <- lowess(x = profile$y, y = profile$value, f=1)$y

profiles <- tidyr::gather(profile, coarseness, values, lw20:lw30)
profiles <- profiles %>% mutate(
  c = readr::parse_number(coarseness)/100
)
p1 <- profiles %>% 
  ggplot(aes(x = y, y = value-values)) + geom_line() +
  theme_bw() +
  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) +
  facet_grid(~c, labeller="label_both")

profc20 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo:160-wv:30-c:20.rds")
profc20$coarseness <- 0.2
profc25 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo:160-wv:30-c:25.rds")
profc30 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo:160-wv:30.rds")
profc30$coarseness <- .30
profc <- bind_rows(profc20, profc25, profc30)
profc <- profc %>% mutate(
  locations = chumbley %>% purrr::map_int(.f = function(x) {
    res <- NULL
    try(res <- x$locations[1], silent = TRUE)
    res
  })
)
profsumm <- read.csv("../data/profiles-summary.csv")
profsumm <- profsumm %>% select(-run_id)
profc <- profc %>% left_join(profsumm, by=c("land2_id"="land_id", "profile2_id"="profile_id"))

p2 <- profc %>% 
  ggplot(aes(x = locations/(length-wo))) +
  geom_histogram(binwidth = 0.025) +
  facet_grid(.~coarseness, labeller="label_both") + 
  theme_bw()

grid.arrange(p1, p2, ncol=1)
```
